Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-05-24 manuscript/revision2/revision2.tex
revision: Address Tom's comment about sign-changing GxE vs antagonistic pleiotropy

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\keywords{allele-by-environment effect variation | antagonistic
pleiotropy | photoperiod | cumulative rainfall | genetic variation}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects, environmental
(e.g., weather) cues, and possibly their interaction. Quantifying these
GxWeather interactions can aid prediction and manipulation of
phenological timings. Here, we map GxWeather effects on phenological
timings in two highly divergent switchgrass (\emph{Panicum virgatum})
populations using repeated plantings of cloned individuals from these
populations at eight sites across the central United States. Some
genetic effects covary with interpretable weather-based cues. For
example, in the Gulf population, 65\% of genetic effects on the timing
of vegetative growth covary with daylength 14 days prior to green-up
date, and 33\% of genetic effects on the timing of flowering covary with
cumulative rainfall in the seven days prior to flowering. However, most
variation in genetic effects across sites is site-specific and can not
be attributed to variation in weather variables alone. We demonstrate
that we can identify genetic variation with GxWeather and assign these
loci to specific weather-based cues or other patterns. Breeding for
particular alleles at the GxWeather loci we identified could change
flowering responsiveness in a photoperiod and rainfall-specific way.
More broadly, our approach proposes a refined, powerful characterization
of genotype-by-environment interactions in any species phenotyped in
multiple environments.
\end{abstract}

@@

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or sign-changing GxE, at the level of
individual loci (23--26), previous empirical work has found limited
evidence of antagonistic pleiotropy (12, 27). However, this work has
been limited by a known statistical bias that reduced detection of
genetic effects that differ in sign (27--29). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

@@

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(Table 1) and additional ways phenological timings might vary by site
(SI Appendix, Table S1, Section SX), then jointly re-estimate genetic
effects on these timings at all eight sites using the set of these
covariance matrices that significantly improved the modeled
log-likelihood when included (SI Appendix, Section SX) (36). We use the
Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
developed by (36), to refine effect size estimates from genome-wide
association (GWAS) conducted on individuals at each of the eight sites.
\emph{mash} allows us to identify and specify multiple covariance
structures among genetic effect estimates across sites, including
structures that represent covariance in weather variables of interest.
Importantly, this method circumvents statistical biases in detecting
genetic effects with the same or opposite signs (37). To confirm our
genetic mapping of GxWeather, we compare our posterior genetic effect
estimates from \emph{mash} to mapping results from an outbred pseudo-F2
cross grown at the same sites. Our analyses allow us to describe the
weather cues and genetic variation affecting phenology in two divergent
natural populations of switchgrass.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the \emph{mash}
model, which can flexibly capture many modes of covariance across
gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
specified three qualitative categories of covariance structures across
gardens: ``canonical'' covariance, with simple patterns of effect size
covariance introduced in the initial \emph{mash} manuscript;
``data-driven'' covariances derived from common patterns of SNP effects
observed in the data, and ``GxWeather'' covariance, estimated from the
covariance of empirical weather patterns at each garden at specific
times before the phenological event (Table 1; SI Appendix, Section S2).
For example, the more similar that the amount of rainfall prior to
flowering is for genotypes A and B between gardens in MI and MO, the
nearer these covariances will be to one. If rainfall amounts are
uncorrelated, covariance will be near zero; if there is rank changing
(e.g.~high rainfall before flowering of A at MO but not MI, and high
rainfall before flowering for B at MI but not MO), this covariance will
be near negative one. We specified multiple covariance structures for
the canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model.

The GxWeather covariance structures allow hypothesis testing of specific
weather variables as cues for the start of vegetative and reproductive
growth. Say that a SNP in FLC controls flowering in a
photoperiod-dependent manner. In that case, the joint estimate of
effects for that SNP could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length at some interval prior to flowering. In our data, we
would infer that the effect of that SNP on flowering was caused by a
response to the environmental cue used to construct the GxWeather
covariance structure with the largest mass. Overall, loadings of genetic
effects on these GxWeather matrices in the model, along with the
posterior effect size estimates, provide information on genome-wide
patterns of SNP-environment interaction.

@@

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither of these
covariance matrices significantly improved the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six \emph{mash} models of
strong effects, the GxWeather covariance matrices captured a minority of
the posterior weights of the strong effects (Figure~\ref{fig-covar}
C,E); in these five models, the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. If the overall
pattern of phenotypic expression differs between a pair of gardens, this
difference may be comprised of SNP effects that differ in sign between
these gardens (effects with antagonistic pleiotropy), SNP effects that
differ in magnitude (effects that are large in one garden or region and
smaller in others), and SNP effects that are indistinguishable in the
two gardens (Figure~\ref{fig-effects} A). We used the local false sign
rate (lfsr), an analogue of the lfdr that establishes confidence in the
effect sign, not the effect's difference from zero, to determine
significance. We required lfsr significance (p \textless{} 0.05) in both
gardens to include effects. This means that our tests for antagonistic
pleiotropy, or a sign change between conditions, carry an equal
statistical burden to those for effects with the same sign.

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} B; SI Appendix, Fig SX). 78.7\% of pairwise
comparisons between North and Texas gardens had a difference in sign,
while only 28.6\% and 0.2\% of North-North or Texas-Texas comparisons
had a difference in sign, respectively (Figure~\ref{fig-effects} B). The
majority of pairwise effects for greenup for the Midwest
(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
same sign, and effects most often differed in magnitude between gardens
within and between the regions (Figure~\ref{fig-effects} B).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects} B). More
effects differed in magnitude between the Texas and North regions than
within these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects}
B), while the majority of effects had no GxE. The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign (Figure~\ref{fig-effects}
B). Most differences in sign were between TX1, the southernmost garden,
and all other gardens (SI Appendix, Fig SX). Similarly, more effects
that differed in magnitude included gardens in the Texas region
(52.3-55.9\%), and most effect pairs in the North region were not
distinguishable (91.5\%).

@@

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites
(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
mapping of flowering as functions of five environmental cues that we
also used as covariance matrices in \emph{mash}, and identified eight
QTL for flowering date, eight QTL for flowering as a function of day
length change two days prior, one QTL for the start of vegetative
growth, and two QTL for vegetative growth as a function of daylength
change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering and
green-up QTL intervals contained at least one SNP significant in at
least one \emph{mash} run at a log10-transformed Bayes Factor
\textgreater{} 2, or in the 1\% tail of significance, whichever was
stricter (SI Appendix, Dataset 8). We also looked for enrichments of
\emph{mash} SNPs in the 1\% tail of significance (the `\emph{mash} 1\%
tail') within each QTL interval. At the 5\% level, three QTL had
enrichments of SNPs in the \emph{mash} 1\% tail. Overall, there were
five significant enrichments (p \textless{} 0.05, hypergeometric test)
of SNPs in the \emph{mash} 1\% tail in the QTL intervals. Thus, we were
able to experimentally support some of our re-estimates of SNP effects
with a QTL mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We can assign genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to agnostic, site-based patterns. We use this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (Figure~\ref{fig-effects} B). This
result supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (23--26), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 38,
39).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on a photoperiod cue. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. Three genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (Figure~\ref{fig-qtl} B).

@@

\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
\bottomrule
\end{tabular}

@@

Last, we characterized the overall patterns of antagonistic pleiotropy
in the set of SNPs where there was pairwise significance of effects at
pairs of gardens. To do this, we used the `get\_GxE' function of the
switchgrassGWAS R package. First, this determines the set of SNPs with
evidence of significant effects in both conditions for all pairs of
conditions using local false sign rates (lfsr) as the significance
criteria. Second, to determine antagonistic pleiotropy, this function
determines if effects significant in both conditions are of opposite
sign.

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect antagonistic pleiotropy. The
lfdr, like other measures of FDR, focuses on if we have enough evidence
to reject the null hypothesis that an effect j is 0, or that there is a
significant effect. Previous studies of antagonistic pleiotropy (e.g.
(39)) have used the lfdr or equivalent statistical tests to detect
antagonistic pleiotropy. These tests were conservative, in that they
required two non-zero effects of different signs, while tests for
differential sensitivity required only one non-zero effect. This
previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy (27, 29), and
sought to reduce it by permutation (28). However, using the lfsr to test
for antagonistic pleiotropy does not undercount occurrences of
antagonistic pleiotropy, as this statistic answers a fundamentally
different question. For each effect j, the lfsr\textsubscript{j} is
defined as the probability that we make an error in the sign of effect j
if we were forced to declare the effect positive or negative (37). Thus,
rather than asking ``Are these two effects different?'' - as we
reasonably expect two effects to be, even if this difference cannot be
measured - the local false sign rate answers a more meaningful question:
Can we be confident in the sign of this effect?

In addition, the get\_GxE function also sets an arbitrary threshold to
count an effect as changing in magnitude between environments, commonly
known as differential sensitivity or a change in amplitude of the
effect. For differential sensitivity, this function determines if
effects significant in both conditions are of the same sign and of a
magnitude (not tested for significance) that differs by a factor of 0.4
or more. The remaining effects that are significant in both conditions
have the same effect sign and similar effect magnitudes and we denote
these effects as having no GxE. The distinction between effects with
different magnitudes is arbitrary but useful to fully characterize how
effects vary across environments to ultimately influence phenotypes. Our
use of the lfsr to determine significance and our specification that SNP
effects must be significant in both conditions to be included means that
our tests for antagonistic pleiotropy carry an equal statistical burden
to those measuring differential sensitivity and effects without GxE.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, rank-changing
genotype-by-environment interaction, photoperiod, cumulative
rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\keywords{allele-by-environment effect variation | rank-changing
genotype-by-environment interaction | photoperiod | cumulative
rainfall | genetic variation}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects, environmental
(e.g., weather) cues, and possibly their interaction. Quantifying these
GxWeather interactions can aid prediction and manipulation of
phenological timings. Here, we map GxWeather effects on phenological
timings in two highly divergent switchgrass (\emph{Panicum virgatum})
populations using repeated plantings of cloned individuals from these
populations at eight sites across the central United States. Some
genetic effects covary with interpretable weather-based cues. For
example, in the Gulf population, 65\% of genetic effects on the timing
of vegetative growth covary with daylength 14 days prior to green-up
date, and 33\% of genetic effects on the timing of flowering covary with
cumulative rainfall in the seven days prior to flowering. However, most
variation in genetic effects is site-specific and can not be attributed
to variation in the studied weather variables. We demonstrate that we
can identify genetic variation with GxWeather and assign these loci to
specific weather-based cues or other patterns. Breeding for particular
alleles at the GxWeather loci we identified could change flowering
responsiveness in a photoperiod or rainfall-specific way. More broadly,
our approach proposes a refined, powerful characterization of
genotype-by-environment interactions in any species phenotyped in
multiple environments.
\end{abstract}

@@

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes (23--26), previous empirical work has found limited evidence of
antagonistic pleiotropy (12, 27). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (27--29). Thus, despite substantial interest in the
frequencies of various forms of GxE, particularly sign-changing GxE,
relative to other forms of GxE remains unknown.

@@

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology for cloned
individuals grown at eight common gardens to many patterns of weather
covariance at these gardens. To do this, we phenotype a diversity panel
of hundreds of switchgrass genotypes from the Midwest and Gulf
subpopulations for the timing of vegetative and reproductive development
at eight common garden locations spanning 17 degrees of latitude. These
gardens cover the majority of the latitudinal and climatic range of
switchgrass and capture the most comprehensive picture to date of the
environmental variation this species encounters. We define multiple ways
phenological timings might covary with weather (Table 1) and additional
ways phenological timings might vary by garden (SI Appendix, Section
S1), then jointly re-estimate genetic effects on these timings at all
eight common gardens using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included (SI
Appendix, Section S2) (36). We use the Bayesian framework \emph{mash}
(multivariate adaptive shrinkage) developed by (36), to refine effect
size estimates from genome-wide association (GWAS) conducted on
individuals at each of the eight gardens. \emph{mash} allows us to
identify and specify multiple covariance structures among genetic effect
estimates across sites, including structures that represent covariance
in weather variables of interest. Importantly, this method circumvents
statistical biases in detecting genetic effects with the same or
opposite signs (37). To confirm our genetic mapping of GxWeather, we
compare our significant posterior genetic effects locations from
\emph{mash} to mapping results from an outbred mapping population grown
at the same sites. Our analyses allow us to describe the weather cues
and genetic variation affecting phenology in two divergent natural
populations of switchgrass across the species' latitudinal range.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects of SNPs across all eight common gardens using the
\emph{mash} model, which can flexibly capture many modes of covariance
across gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
specified three qualitative categories of covariance structures across
gardens: ``canonical'' covariance, with simple patterns of effect size
covariance introduced in the initial \emph{mash} manuscript;
``data-driven'' covariances derived from common patterns of SNP effects
observed in the data, and ``GxWeather'' covariance, estimated from the
covariance of empirical weather patterns at each garden at specific
times before the phenological event (Table 1; SI Appendix, Section S2).
For example, the more similar that the amount of rainfall prior to
flowering is for genotypes A and B between gardens in MI and MO, the
nearer these covariances will be to one. If rainfall amounts are
uncorrelated, covariance will be near zero; if there is rank changing
(e.g.~high rainfall before flowering of A at MO but not MI, and high
rainfall before flowering for B at MI but not MO), this covariance will
be near negative one. We specified multiple covariance structures for
the canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model.

The GxWeather covariance structures allow hypothesis testing of specific
weather variables as cues for the start of vegetative and reproductive
growth. Say that a SNP in the gene \emph{Flowering Locus C} (FLC), a
well-known flowering time regulator, controls flowering in a
photoperiod-dependent manner. In that case, the joint estimate of
effects for that SNP could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length at some interval prior to flowering. In our data, we
would infer that the effect of that SNP on flowering was caused by a
response to the environmental cue used to construct the GxWeather
covariance structure with the largest mass. Overall, loadings of genetic
effects on these GxWeather matrices in the model, along with the
posterior effect size estimates, provide information on genome-wide
patterns of SNP-environment interaction.

@@

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither of these
covariance matrices significantly improved the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six \emph{mash} models of
strong effects, the GxWeather covariance matrices captured a minority of
the posterior weights of the strong effects (Figure~\ref{fig-covar}
C,E); in these five models, the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and single effect matrices with garden-specific effects (SI
Appendix, Section S1).

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. If the overall
pattern of phenotypic expression differs between a pair of gardens, this
difference may be comprised of SNP effects that differ in sign between
these gardens (effects with rank-changing GxE between gardens), SNP
effects that differ in magnitude (effects that are large in one garden
or region and smaller in others), and SNP effects that are
indistinguishable in the two gardens (Figure~\ref{fig-effects} A). We
used the local false sign rate (lfsr), an analogue of the lfdr that
establishes confidence in the effect sign, not the effect's difference
from zero, to determine significance. We required lfsr significance (p
\textless{} 0.05) in both gardens to include effects. This means that
our tests for a sign change between gardens carry an equal statistical
burden to those for effects with the same sign.

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited a difference in effect sign, between pairs of
Texas and North gardens (Figure~\ref{fig-effects} B; SI Appendix, Fig
SX). 78.7\% of pairwise comparisons between North and Texas gardens had
a difference in sign, while only 28.6\% and 0.2\% of North-North or
Texas-Texas comparisons had a difference in sign, respectively
(Figure~\ref{fig-effects} B). The majority of pairwise effects for
greenup for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
subpopulations were the same sign, and effects most often differed in
magnitude between gardens within and between the regions
(Figure~\ref{fig-effects} B).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited a difference in effect sign within or between regions
(Figure~\ref{fig-effects} B). More effects differed in magnitude between
the Texas and North regions than within these regions (42.7\% vs
\textless20\%; Figure~\ref{fig-effects} B), while the majority of
effects had no GxE. The Midwest population had relatively few
significant effects for flowering, but a large proportion of these
differed in sign between Texas and North regions (42.7\%) or within the
North region (65.4\%). Finally, in Both subpopulations, less than 20\%
of pairwise effects differed in sign (Figure~\ref{fig-effects} B). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (SI Appendix, Fig SX). Similarly, more effects that
differed in magnitude included gardens in the Texas region
(52.3-55.9\%), and most effect pairs in the North region were not
distinguishable (91.5\%).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering and
green-up QTL intervals contained at least one SNP significant in at
least one \emph{mash} run at a log10-transformed Bayes Factor
\textgreater{} 2, or in the 1\% tail of significance, whichever was
stricter (SI Appendix, Dataset 8). We also looked for enrichments of
\emph{mash} SNPs in the 1\% tail of significance (the `\emph{mash} 1\%
tail') within each QTL interval. At the 5\% level, three QTL had
enrichments of SNPs in the \emph{mash} 1\% tail. Overall, there were
five significant enrichments (p \textless{} 0.05, hypergeometric test)
of SNPs in the \emph{mash} 1\% tail in the QTL intervals. Thus, we were
able to experimentally support some of our re-estimates of SNP effects
with a QTL mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We can assign genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to unmeasured, site-based patterns. We use this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed many alleles with sign changes in their effects
between the Texas and North gardens (Figure~\ref{fig-effects} B). As
phenological timings are major components of plant fitness, this result
supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (23--26); it
expands the genetics of local adaptation research to more than two field
sites and to a wider range of genetic variation by using effects from
GWAS in common gardens located throughout the species' range to
determine that rank-changing GxE for phenological timings is common in
small genomic regions (12, 38, 39).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related genetic variation, in that
flowering timing covaries with a day length change signal two days
before flowering occurs. In contrast, the Gulf subpopulation does not
have genetic variation in flowering that covaries with a photoperiod
cue. Instead, the Gulf subpopulation has genetic variation in flowering
that covaries with the rainfall that occurs in the week prior to
flowering. Three genomic regions affecting flowering that we
re-estimated across all eight sites were also supported by QTL from an
independent mapping population at these sites (Figure~\ref{fig-qtl} B).

@@

\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
\bottomrule
\end{tabular}

@@

Last, we characterized the overall patterns of GxE in the set of SNPs
where there was pairwise significance of effects at pairs of gardens. To
do this, we used the `get\_GxE' function of the switchgrassGWAS R
package. First, this determines the set of SNPs with evidence of
significant effects in both conditions for all pairs of conditions using
local false sign rates (lfsr) as the significance criteria. Then, this
function determines if effects significant in both conditions are of
opposite sign.

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect alleles directly contributing
to rank-changing GxE between environments. The lfdr, like other measures
of FDR, focuses on if we have enough evidence to reject the null
hypothesis that an effect j is 0, or that there is a significant effect.
Previous studies of antagonistic pleiotropy (e.g. (39)) have used the
lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
These tests were conservative, in that they required two non-zero
effects of different signs, while tests for differential sensitivity
required only one non-zero effect. This previous work recognized that
this testing bias could lead to undercounting occurrences of
antagonistic pleiotropy (27, 29), and sought to reduce it by permutation
(28). However, using the lfsr to test for allelic effects that differ in
sign does not undercount these occurences, as this statistic answers a
fundamentally different question. For each effect j, the
lfsr\textsubscript{j} is defined as the probability that we make an
error in the sign of effect j if we were forced to declare the effect
positive or negative (37). Thus, rather than asking ``Are these two
effects different?'' - as we reasonably expect two effects to be, even
if this difference cannot be measured - the local false sign rate
answers a more meaningful question: Can we be confident in the sign of
this effect?

In addition, the get\_GxE function also sets an arbitrary threshold to
count an effect as changing in magnitude between environments, commonly
known as differential sensitivity or a change in amplitude of the
effect. For differential sensitivity, this function determines if
effects significant in both conditions are of the same sign and of a
magnitude (not tested for significance) that differs by a factor of 0.4
or more. The remaining effects that are significant in both conditions
have the same effect sign and similar effect magnitudes and we denote
these effects as having no GxE. The distinction between effects with
different magnitudes is arbitrary but useful to fully characterize how
effects vary across environments to ultimately influence phenotypes. Our
use of the lfsr to determine significance and our specification that SNP
effects must be significant in both conditions to be included means that
our tests for alleles with rank-changing GxE carry an equal statistical
burden to those measuring differential sensitivity and effects without
GxE.
